---
title: "WEEK 34 EPIDEMIOLOGICAL BULLETIN (19th – 25th Aug, 2024)"
format:
  docx:
    reference-doc: template.docx
    toc: false
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
# ============================================================
# AUTOMATED BULLETIN PIPELINE
# ============================================================

suppressPackageStartupMessages({
  library(tidyverse)
  library(readr)
  library(knitr)
  library(rprojroot)
})

ROOT <- rprojroot::find_root(rprojroot::is_rstudio_project)
SCRIPT_DIR <- file.path(ROOT, "scripts")
OUT_DIR <- file.path(ROOT, "outputs")

source(file.path(SCRIPT_DIR, "01_load_and_validate_data.R"))
source(file.path(SCRIPT_DIR, "02_week34_tables_and_text.R"))
source(file.path(SCRIPT_DIR, "03_plots_and_maps.R"))

priority_table <- read_csv(file.path(OUT_DIR, "priority_table_week34.csv"), show_col_types = FALSE)
narratives <- readLines(file.path(OUT_DIR, "week34_narratives.txt"))
measles_ind <- readLines(file.path(OUT_DIR, "measles_lab_indicators.txt"))
md_total <- readLines(file.path(OUT_DIR, "maternal_deaths_cumulative_total.txt"))

img_measles  <- file.path(OUT_DIR, "measles_lab_by_province.png")
img_md_cause <- file.path(OUT_DIR, "maternal_deaths_causes_week34.png")
img_md_map   <- file.path(OUT_DIR, "maternal_deaths_cumulative_map.png")
```

# 1. Summary

## Key highlights for decision-makers

- **Measles:** Suspected cases were reported in multiple provinces. Strengthen active case search, specimen referral, and rapid response in high-reporting areas.
- **Anthrax:** Suspected cases were concentrated in Southern Province, indicating continued zoonotic risk. Strengthen One Health coordination and laboratory confirmation.
- **Malaria:** Malaria remains the leading contributor to weekly morbidity. Prioritize hotspot districts and ensure commodity availability.

# 2. Immediately notifiable diseases and events

```{r}
cat(paste(narratives[1:4], collapse = "\n"))
```

## Interpretation and recommendations

Prompt verification and investigation of suspected cases is recommended, particularly where clustering is observed. Provinces should ensure timely reporting and adherence to standard case definitions.

# 3. Other diseases and events

```{r}
cat(paste(narratives[6:8], collapse = "\n"))
```

## Interpretation and recommendations

Routine high burden conditions require sustained surveillance, health promotion, and facility readiness. Unusual increases should be investigated to rule out outbreaks.

# 4. Summary report: priority diseases, conditions and events

```{r}
library(flextable)
library(officer)

ft <- flextable(priority_table)

ft <- ft |>
  theme_vanilla() |>
  fontsize(size = 10, part = "all") |>
  bold(part = "header") |>
  bg(part = "header", bg = "#F39C34") |> 
  color(part = "header", color = "white") |>
  bg(j = 1, part = "body", bg = "#FCE5CD") |>
  align(align = "center", part = "all") |>
  align(j = 1, align = "left", part = "all") |>
  autofit()

ft

```

# 5. Summary of VPD surveillance indicators

```{r}
cat(paste(measles_ind, collapse = "\n"))
```

```{r}
knitr::include_graphics(img_measles)
```

## Interpretation and recommendations

Laboratory findings indicate ongoing transmission risk in some provinces. Improve specimen quality, cold chain management, and turnaround time.

# 6. Maternal deaths

```{r}
knitr::include_graphics(img_md_cause)
```

```{r}
knitr::include_graphics(img_md_map)
```

```{r}
cat(md_total)
```

## Interpretation and recommendations

Strengthen emergency obstetric care readiness, referral systems, and implementation of MPDSR recommendations, particularly in high-burden provinces.

# 7. Recommendations to improve this epidemiological bulletin

1. Include week-on-week comparisons.
2. Add outbreak thresholds.
3. Standardize visual presentation.
4. Include data quality notes.
5. Add an actions taken / next steps section.

---
Reproducibility note: All outputs are generated automatically from scripts in the scripts folder.
